---
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
#### Global Chunk Options ####
knitr::opts_chunk$set(# Code output options
                      echo=TRUE,message=FALSE,warning=FALSE,
                      # Figure output options
                      fig.height=6,fig.width=8,
                      # Figure saving options
                      fig.path=("Figs/"),dev=c("png"),dpi=200)
options(tinytex.verbose = TRUE)
options(allow_html_in_all_outputs=TRUE)
library(tidyverse)
setwd("C:\\Users\\lykha\\OneDrive\\Documents\\1_PhD_Spring_2020\\Spatial-Temporal Analysis\\Project\\")
```

I will use Texas_HF data in this analysis. Data contains the lung cancer cases and population stratified by gender and race in each of the Texas_HF counties from 2005 to 2016. 

# Prepare spatio-temporal data

```{r readData, include=TRUE}
library(rgdal)
library(sf)
map <- readOGR("C:\\Users\\lykha\\OneDrive\\Documents\\1_PhD_Spring_2020\\Spatial-Temporal Analysis\\Project\\Data\\HF-Data\\Heart_Failure_Texas_2014to2016", "Heart_Failure_Hospitalization_Rate_per_1000_Medicare_Beneficiaries_65plus_All_Races_Ethnicities_Both_Genders_2014to2016", verbose = F)
plot(map)
```
```{r}
Texas_HF_2 <- read_csv("C:\\Users\\lykha\\OneDrive\\Documents\\1_PhD_Spring_2020\\Spatial-Temporal Analysis\\Project\\Data\\HF-Data\\Raw_HF_Data\\New_extracted\\HF_2006-16_gender_race_rate.csv")
Texas_HF_2$Y[Texas_HF_2$Y=="-1"] <- 0
Texas_HF_2
```

Variables:
dm_prev_adj,
ob_prev_adj,
ltpia_prev_adj
perc_aian
perc_api
perc_black
perc_white
perc_hisp
perc_65up


```{r}
Texas_HF = merge(map@data[,c(2,5)], Texas_HF_2, by.x="cnty_fips", by.y="cnty_fips")
Texas_HF
```

```{r}
d <- aggregate(x=Texas_HF$Y, by=list(county=Texas_HF$cnty_name, year=Texas_HF$year), FUN=sum)
names(d) <- c("county", "year", "Y")
d
```

```{r expectedData, include=TRUE}
### sum over all cases and population for each strata
dat_strat <- Texas_HF %>%  summarise(y_total = sum(Y), N_total = sum(total_pop))
dat_strat
### reference probs
q <- with(dat_strat, y_total/N_total)
q
```

### calculate expected cases

```{r}
d2 <- Texas_HF %>% group_by(cnty_name, year) %>%  summarise(y_total = sum(Y), N_total = sum(total_pop))
d2$q <- rep(q, nrow(d))
d2$E_strat <- d2$N_total * d2$q
E_dat <- d2 %>% group_by(cnty_name, year) %>%  summarise(E = sum(E_strat))
d <- merge(d, E_dat, by.x=c("county", "year"), by.y=c("cnty_name", "year"))
d
```

### Calculate the SIR and remove NA values
```{r SIR, include=TRUE}
d$SIR <- with(d, Y/E)
d
d$Y[which(d$Y <0)] <- NA
```

```{r addMap, include=TRUE}
dw <- reshape(d, timevar="year", idvar="county", direction="wide")
dw[1:3,]
map@data[1:2,]
map <- merge(map, dw, by.x="cnty_name", by.y="county")
map@data
```

### Drawing the map of SIR

```{r mappingSIR}
map_sf <- st_as_sf(map)
library(tidyr)
map_sf <- gather(map_sf, year, SIR, paste0("SIR.", c("2005-2007", "2006-2008", "2007-2009", "2008-2010", "2009-2011", "2010-2012", "2011-2013", "2012-2014", "2013-2015", "2014-2016")))
map_sf$year <- substring(map_sf$year, 5, 13)

library(ggplot2)
ggplot(map_sf) + geom_sf(aes(fill=SIR)) + facet_wrap( ~ year, dir="h", ncol=7) + 
  ggtitle("SIR") + theme_bw() + 
  theme(axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks =element_blank()) + 
  scale_fill_gradient2(midpoint=1, low="blue", mid="white", high="red")
```

### Time trend plots:

```{r, timeplot}
g <- ggplot(d, aes(x=year, y=SIR, group=county, color=county)) + 
  geom_line() + geom_point(size=2) + theme_bw()
print(g)

g <- g+ theme(legend.position = "none")
library(gghighlight)
g + gghighlight(county == "Harris")
print(g)
```


```{r plotly}
library(plotly)
ggplotly(g)
```

# Modeling
We estimate the relative risk of lung cancer for each Texas_HF county and year, using the Bernardinelli model (Bernardinelli et al, 1995). 

```{r model-nei}
### build neighborhood matrix
library(INLA)
library(spdep)
nb <- poly2nb(map)
head(nb)
nb2INLA("Texas_HF_map.adj", nb)
g <- inla.read.graph(filename="Texas_HF_map.adj")
```
```{r}
d2 <- merge(d, Texas_HF[, c(2,7,8,9,10,11, 13, 14, 15, 16, 18, 19)], by.x=c("county", "Y", "year"), by.y=c("cnty_name", "Y", "year"))
d2
```


```{r model-1}
d2$id.s <- as.numeric(d2$county)
d2$id.k <- d2$id.u
d2$id.area <- d2$id.u
d2$id.time <- 1 + as.integer(substring(d2$year, 1, 4)) - min(as.integer(substring(d2$year, 1, 4)))
d2$id.time2 <- d2$id.time
d2$id.timespace <- 1:nrow(d2)

f <- Y ~ f(id.s, model="besag", graph=g) + f(id.k, model="iid") + f(id.time, model="rw1") + f(id.time2, model="iid") + f(id.timespace, model="iid") + dm_prev_adj + ob_prev_adj + ltpia_prev_adj + perc_black + perc_white + perc_hisp + perc_65up + povpct
```

### Full Model:
```{r}
mod <- inla(f, family="poisson", data=d2, E=E,  control.predictor = list(compute=TRUE), control.compute = list(cpo=TRUE, dic=TRUE, waic=TRUE))
summary(mod)
d2$RR <- mod2$summary.fitted.values[, "mean"]
d2$LL <- mod$summary.fitted.values[, "0.025quant"]
d2$UL <- mod$summary.fitted.values[, "0.975quant"]
```
### Dropped Model:
I dropped all the insignificant variables (keeping only Poverty variable) and then compare the new with the original model. The new WAIC is 8797.23 and new DIC is 9195.50, which are slightly higher than the original criterions. Therefore, I decided to choose the original model as my final model. 
```{r}
f2 <- Y ~ f(id.s, model="besag", graph=g) + f(id.k, model="iid") + f(id.time, model="rw1") + f(id.time2, model="iid") + f(id.timespace, model="iid") + + povpct
mod2 <- inla(f2, family="poisson", data=d2, E=E,  control.predictor = list(compute=TRUE), control.compute = list(cpo=TRUE, dic=TRUE, waic=TRUE))

summary(mod2)
```


### Drawing the map of RR from full model:
```{r modelmapping}
map_sf <- merge(map_sf, d2, by.x=c("cnty_name", "year"), by.y=c("county", "year"))
ggplot(map_sf) + geom_sf(aes(fill=RR)) + facet_wrap( ~ year, dir="h", ncol=7) + 
  ggtitle("RR") + theme_bw() + 
  theme(axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks =element_blank()) + 
  scale_fill_gradient2(midpoint=1, low="blue", mid="white", high="red")

#myrange <- range(d$RR, d$RR2)
#plot(RR2 ~ RR, xlab="Model 1", Ylab="Model 2", data=d, xlim=myrange, ylim=myrange)
#abline(a=0, b=1, col=2)
```


# Animation
Create animated maps showing the relative risks for each year. To save the animation. use anim\_save() function whicy by default saves a file of type gif.

setwd("C:\\Users\\lykha\\OneDrive\\Documents\\1_PhD_Spring_2020\\Spatial-Temporal Analysis\\Project\\Output_video")
install.packages("transformr")
library(gganimate)
ggplot(map_sf) + geom_sf(aes(fill=RR)) + theme_bw() + 
 theme(axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks =element_blank()) + 
 scale_fill_gradient2(midpoint=1, low="blue", mid="white", high="red") + 
 transition_time(as.integer(substring(year, 1, 4))) + 
 labs(title="Year: {round(frame_time, 0)}")

Exceedance probabilities
We can also calculate the probabilities of relative risk estimates being greater than a given threshold value. These probabilities are called exceedance probabilities and are useful to assess unusual elevation of disease risk. The probability that the relative risk of area 

### Exceedance Probabilities:
I eventually calculated the probabilities of relative risk estimates being greater than 2, which indicates as high risk for HF hospitalization. These probabilities are called exceedance probabilities and are useful to assess unusual elevation of disease risk. The probability that the relative risk of area i is higher than a value 2 is  P(θi>2), which can be calculated as 1−P(θi≤2).
```{r}
exc <- sapply(mod2$marginals.fitted.values, FUN=function(marg) {1 - inla.pmarginal(q=2, marginal = marg)})
d2$Exc <- exc
```

```{r}
d2
```

### Drawing the map of exceedance probabilities:
```{r}
map_sf <- merge(map_sf, d2, by.x=c("cnty_name", "year"), by.y=c("county", "year"))
ggplot(map_sf) + geom_sf(aes(fill=Exc)) + facet_wrap( ~ year, dir="h", ncol=7) + 
  ggtitle("P(RR>2)") + theme_bw() + 
  theme(axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks =element_blank()) + 
  scale_fill_gradient2(midpoint=1, low="white", mid="red", high="darkred") +labs(fill="P(RR>2)")
```




